Intentional release of native species undermines ecological stability

### 2022-09-28 rmd/supporting_info.Rmd
Update supporting_info.Rmd

---
output:
  bookdown::pdf_document2:
    number_sections: false
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
mainfont: "Times New Roman"
fontsize: 11pt
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{amsmath,fancyhdr, caption}
  - \captionsetup{width=1.3\linewidth, labelfont={bf}}
  - \pagestyle{fancy}
bibliography: reference.bib
csl: https://www.zotero.org/styles/pnas
---

@@

```{=tex}
\begin{align}
\varepsilon_{i,t}^{\text{state}}=\sum_{d}^{D}\zeta_{t,d}\delta_{d,i} + I_{ij}\epsilon_{i,t}
\end{align}
```
where $D$ is the number of latent factors, $\epsilon_{i,t} \sim \text{Normal}(0, \sigma_{\epsilon,i}^2)$, and $I_{ij}$ is the identity matrix ($I_{ii}=1$ and $I_{ij, i \ne j} = 0$). With this formulation, the matrix $\pmb{\Omega_{\varepsilon}}$ can be described as $\pmb{\Omega_{\varepsilon}}=\pmb{\Delta}^T\pmb{\Delta} + \text{diag}(\sigma_{\epsilon,i}^2)$, where $\pmb{\Delta}$ is the matrix of factor loadings $\delta_{d,i}$. This model structure has a merit of reduced number of parameters while capturing the correlated nature of environmental noises [@ovaskainenHowAreSpecies2017]. We set $D=2$ in our model. We reported the estimates of $\alpha_{ij}$ so the values are comparable to those used in the theoretical analysis.

@@

## Table S1 Simulation parameter

Parameter values used in the whole community simulation.

```{r table_s1}

# Table S1
suppressMessages(source(here::here("code/table_param_set.R")))
knitr::kable(df_param, format = "markdown")

@@

## Table S2 Observed species

Fish species found in the study watersheds.

```{r table_s2}

# Table S2
suppressMessages(source(here::here("code/table_species_list.R")))

knitr::kable(df_sp_list,
             format = "markdown")
```

@@

```{r table_s3}

# Table S6
suppressMessages(source(here::here("code/table_prior.R")))

@@

## Parameter estimates of the Bayesian state-space model

```{r table_s7}

# Table S7
suppressMessages(source(here::here("code/table_ssm.R")))

@@

# Table S8-10
## load regression tables
suppressMessages(source(here::here("code/table_reg.R")))
j <- 7

@@

## common text
table_legend <- "Median estimates and standard errors (SEs) of regression coefficients are shown. Pr(> 0) and Pr(< 0) represent the proportion of positive and negative estimates in MCMC samples, respectively (i.e., posterior probability)."

table_reg_chunk <- knit_chunk <- NULL

@@

  knit_chunk <- paste0("## Table S", i + j, " Parameter estimates for the regression model (", c_text[i], ") \n",
                       table_legend,
                       "\n```{r table_s", i + j,"}
                       \n\n
                       kable(list_reg[[",i,"]],
                             align = c(rep('l', 3), rep('c', 4)),
                             format = 'markdown')
                       \n\n
                       ```\n",
                       "\n\\pagebreak\n")

@@

j <- 0
figure_sim_chunk <- knit_chunk <- NULL

@@

  knit_chunk <- paste0("## Figure S", i + j,
                       " Theoretical prediction (",
                       "$r_1$ = `r df_param$r1[",i,"]`, ",
                       "$K$ = `r df_param$k[",i,"]`) \n",

                       "\n```{r figure_s", i + j,", fig.width = 10, fig.height = 7}
                       \n\n
                       list_g_theory[[",i,"]]
                       \n\n
                       ```\n",

                       "Theoretical predictions for the release effect on community dynamics. Rows represent different response variables, and columns show distinct simulation scenarios with different strength of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Other parameters are: ",
                       "number of species $S$ = `r df_param$n_species[",i,"]`; ",
                       "intrinsic growth rate of an enhanced species $r_1$ = `r df_param$r1[",i,"]`; ",
                       "maximum intrinsic growth rate of unenhanced species $r_{max}$ = `r df_param$r_max[",i,"]`; ",
                        "environmental stochasticity $\\sigma_{\\varepsilon}$ = `r df_param$sd_env[",i,"]`; ",
                        "carrying capacity $K$ = `r df_param$k[",i,"]`.",
                        "\n\\pagebreak\n")

@@

```{r figure_dyns}

# Figure S9-11
## load figures
suppressMessages(source(here::here("code/si_figure_obs_dyns.R")))
j <- 8

@@

figure_dyns_chunk <- knit_chunk <- NULL

@@

  knit_chunk <- paste0("## Figure S", i + j,
                       " Temporal dynamics of stream fish communities (`r c_text[",i,"]`) \n",

                       "\n```{r figure_s", i + j,", fig.width = 11, fig.height = 11}
                       \n\n
                       list_g_dyns[[",i,"]]
                       \n\n
                       ```\n",
                       "Temporal dynamics of stream fish communities (`r c_text[",i,"]`) in Hokkaido, Japan. Dots represent observed density, and solid lines are the predicted values of the Bayesian state-space model. Panels correspond to individual watersheds, and colors distinguish sampling sites within a watershed.",
                        "\n\\pagebreak\n")

@@

## Figure S12 Environmental variables

```{r figure_s12, fig.width=11, fig.height=8}

@@

Distribution of environmental variables in the protected watershed. Note that the number of fish released and ocean productivity (chlorophyll *a*) are measured at the watershed level while others are measured at the site level.

@@

## Figure S13 Correlation plot

```{r figure_s13, fig.width=11, fig.height=11}

@@

Correlation plot among environmental variables in the protected watersheds. Numbers indicate Spearman's rank correlations.

>>>>

---
output:
  bookdown::pdf_document2:
    fig_caption: true
    number_sections: false
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
mainfont: "Times New Roman"
fontsize: 11pt
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
header-includes:
  - \pagenumbering{gobble}
  - \usepackage{amsmath,fancyhdr, caption}
  - \captionsetup{width=1.3\linewidth, labelfont={bf}}
  - \pagestyle{fancy}
  - \fancyhead[L]{}
  - \fancyhead[C]{}
  - \setlength{\headheight}{13.7pt}
bibliography: reference.bib
csl: https://www.zotero.org/styles/pnas
---

@@

```{=tex}
\begin{align}
\varepsilon_{i,t}^{\text{state}}=\sum_{d}^{D}\zeta_{t,d}\delta_{d,i} + I_{ij}\epsilon_{i,t}
\end{align}
```
where $D$ is the number of latent factors, $\epsilon_{i,t} \sim \text{Normal}(0, \sigma_{\epsilon,i}^2)$, and $I_{ij}$ is the identity matrix ($I_{ii}=1$ and $I_{ij, i \ne j} = 0$). With this formulation, the matrix $\pmb{\Omega_{\varepsilon}}$ can be described as $\pmb{\Omega_{\varepsilon}}=\pmb{\Delta}^T\pmb{\Delta} + \text{diag}(\sigma_{\epsilon,i}^2)$, where $\pmb{\Delta}$ is the matrix of factor loadings $\delta_{d,i}$. This model structure has a merit of reduced number of parameters while capturing the correlated nature of species responses to environmental fluctuations [@ovaskainenHowAreSpecies2017]. We set $D=2$ in our model. We reported the estimates of $\alpha_{ij}$ so the values are comparable to those used in the theoretical analysis.

@@

## Simulation parameter

```{r tab_para}

suppressMessages(source(here::here("code/table_param_set.R")))
knitr::kable(df_param,
             format = "markdown",
             caption = "Parameter values used in the whole community simulation.")

@@

## Observed species

```{r tab_splist}

suppressMessages(source(here::here("code/table_species_list.R")))

knitr::kable(df_sp_list,
             format = "markdown",
             caption = "Fish species found in the study watersheds.")
```

@@

```{r tab_prior}

suppressMessages(source(here::here("code/table_prior.R")))

@@

## Parameter estimates for the Bayesian state-space model

```{r tab_ssm}

suppressMessages(source(here::here("code/table_ssm.R")))

@@

## load regression tables
suppressMessages(source(here::here("code/table_reg.R")))
j <- 7

@@

table_reg_chunk <- knit_chunk <- table_legend <- NULL

@@

  table_legend[i] <- paste0("Parameter estimates for the regression model (", c_text[i], "). Median estimates and standard errors (SEs) of regression coefficients are shown. Pr(> 0) and Pr(< 0) represent the proportion of positive and negative estimates in MCMC samples, respectively (i.e., posterior probability).")

  knit_chunk <- paste0("## Parameter estimates for the regression model (", c_text[i], ")",
                       "\n```{r tab_reg", i, "}
                       \n\n
                       kable(list_reg[[",i,"]],
                             align = c(rep('l', 2), 'r', rep('c', 3)),
                             format = 'markdown',
                             caption = table_legend[",i,"])
                       \n\n
                       ```\n",
                       "\n\\pagebreak\n")

@@

figure_sim_chunk <- knit_chunk <- cap <- NULL

@@

  cap[i] <- paste0("Theoretical predictions for the release effect on community dynamics. Rows represent different response variables, and columns show distinct simulation scenarios with different strength of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Other parameters are: ",
                       "number of species $S$ = ", df_param$n_species[i], "; intrinsic growth rate of an enhanced species $r_1$ = ", df_param$r1[i], "; maximum intrinsic growth rate of unenhanced species $r_{max}$ = ", df_param$r_max[i], "; environmental stochasticity $\\sigma_{\\varepsilon}$ = ", df_param$sd_env[i], "; carrying capacity $K$ = ", df_param$k[i], ".")

  knit_chunk <- paste0("## Theoretical prediction (",
                       "$r_1$ = `r df_param$r1[",i,"]`, ",
                       "$K$ = `r df_param$k[",i,"]`) \n",
                       "\n```{r fig_sim", i,", fig.width = 10, fig.height = 7, fig.cap = cap[", i, "]}
                       \n\n
                       list_g_theory[[",i,"]]
                       \n\n
                       ```\n",
                        "\n\\pagebreak\n")

@@

```{r fig_dyns}

# Figure S9-11
## load figures
suppressMessages(source(here::here("code/si_figure_obs_dyns.R")))

@@

figure_dyns_chunk <- knit_chunk <- cap <- NULL

@@

  cap[i] <- paste0("Temporal dynamics of stream fish communities (", c_text[i], ") in Hokkaido, Japan. Dots represent observed density, and solid lines are the predicted values of the Bayesian state-space model. Panels correspond to individual watersheds, and colors distinguish sampling sites within a watershed.")

  knit_chunk <- paste0(
    "## Temporal dynamics of stream fish communities (`r c_text[",i,"]`)\n",
    "\n```{r fig_d", i,", fig.width = 10, fig.height = 7, fig.cap=cap[", i ,"]}
    \n\n
    list_g_dyns[[",i,"]]
    \n\n
    ```\n",
    "\n\\pagebreak\n")

@@

## Environmental variables

```{r fig_env, fig.width=11, fig.height=8, fig.cap="Distribution of environmental variables in the protected watershed. Note that the number of fish released and ocean productivity (chlorophyll a) are measured at the watershed level while others are measured at the site level."}

@@

@@

## Correlation plot

```{r fig_cor, fig.width=11, fig.height=11, fig.cap="Correlation plot among environmental variables in the protected watersheds. Numbers indicate Spearman's rank correlations."}

@@
